Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-08-31 paper_comparison/paper_comparison/paper_comparison.tex
update the writing and code in main doc

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\begin{quote}
\emph{``Since all models are wrong the scientist must be alert to what
is importantly wrong.''} \citep{REF}
\end{quote}

Diagnosing a model is the key to determining whether there is anything
importantly wrong. For linear regression analysis, it is typical to
interrogate the residuals. Residuals summarise what is not captured by
the model, and thus provide the capacity to identify what might be
wrong. There are many ways that residuals are assessed.

Residuals might be plotted, as a histogram or normal probability plot to
examine the distribution. Is it symmetric, and unimodal? Then all is
good. If the distribution is skewed or bimodal or multimodal, there are
outliers, there is cause for concern. It if also possible to conduct a
goodness of fit test, such as a Shapiro-Wilk Normality test \citep{REF}.

It is also recommended to plot residuals against predicted values, and
also each of the explanatory variables. If there is no obvious pattern,
then all is good. It is a very difficult task for a human judge, though
to make a decision that there's nothing there. It is especially common,
particularly among new data analysts to report patterns when an
experienced data analyst might quickly conclude that there are none.
Generally, one looks for departures from nothingness like nonlinear
dependency or heteroskedasticity. It is also possible to conduct a
hypothesis test for non-linear dependence \citep{RESET}, and also use a
Breusch-Pagan test \citep{REF} for heteroskedasticity.

There is an abundance of literature describing appropriate diagnostic
methods for linear regression: Cook and Weisberg, \ldots{} \citep{REF}.
The diligent reader of these sage writings will also notice sentences
that express sentiments like \emph{based on their experience,
statistical tests are not widely used in regression diagnostics. The
same or even larger amount of information can be provided by diagnostic
plots than the corresponding tests in most empirical studies.} There is
a common guidance by experts that plots are the best for diagnosing
model fits.

This is curious, and investigating why this might be common advice is
the subject of this paper. The paper is structured as follows. The next
background section describes the commonly used diagnostic plots, the
types of departures that one expects to detect, and describes a formal
process for reading residual plots, called visual inference, that can
avoid the concerns about subjectiveness of human readers. Section XXX
describes the experimental setup to enable a comparison between
decisopns made by formal hypothesis testing, and how humans would read
diagnostic plots. The results are reported in Section XXX. We finish
with a discussion on future work, in particular how the responsibility
for residual plot reading might be passed on to computer vision.

@@

\hypertarget{hypothesis-testing}{%
\subsection{Hypothesis testing}\label{hypothesis-testing}}

Other than checking diagnostic plots, analysts may perform formal
hypothesis testing for detecting model defects. Depending on the
alternative hypothesis that is focused on, a variety of tests can be
applied. For example, the presence of heteroskedasticity can usually be
tested by applying the White test
\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
test \citep{breusch_simple_1979}, which are both derived from the
Lagrange multiplier test \citep{silvey1959lagrangian} principle that
relies on the asymptotic properties of the null distribution. For
testing non-linearity, one may apply the F-test to examine the
significance of specific polynomial and non-linear forms of the
regressors, or the significance of proxy variables as in the Ramsey
Regression Equation Specification Error Test (RESET)
\citep{ramsey_tests_1969}.

As discussed in \citet{cook1982residuals}, most residual-based tests for
a particular type of departure from model assumptions are sensitive to
other types of departures. It is likely the null hypothesis is correctly
rejected but for the wrong reason, which is known as the ``Type III
error''. Additionally, outliers will often incorrectly trigger the
rejection of the null hypothesis despite the residuals are well-behaved
\citep{cook_applied_1999}. This can be largely avoided in diagnostic
plots as experienced analysts can evaluate the acceptability of
assumptions flexibly, even in the presence of outliers.
\citet{montgomery1982introduction} suggested that based on their
experience, statistical tests are not widely used in regression
diagnostics. The same or even larger amount of information can be
provided by diagnostic plots than the corresponding tests in most
empirical studies. Not to mention, it is almost impossible to have an
exactly correctly specified model in reality. There is a well-known
aphorism in statistics stated by George Box - ``All models are wrong,
but some are useful''. This indicates proper hypothesis tests will
always reject the null hypothesis as long as the sample size is large
enough. The outcome ``Not reject'' can be interpreted as either ``effect
size is small'' or ``sample size is small''. and the outcome ``reject''
doesn't inform us whether and how much the model defects are of actual
consequence to the inference and prediction. But still, the
effectiveness of statistical tests shall not be disrespected.
Statistical tests have a chance to provide analysts with unique
information. In situations where no suitable diagnostic plots can be
found for a particular violation of the assumptions, or excessive
diagnostic plots are needed to be checked, one will have no choice but
to fall back on statistical tests if there exists any. A good regression
diagnostic practice should be a balanced combination of both methods.

@@

In overall, there were a total of 3200 lineup evaluations made by 160
subjects in experiment I, where 320 lineup evaluations were attention
checks and were not used in the following analysis. The collated dataset
is provided in \texttt{polynomials} of the \texttt{visage} \texttt{R}
package.

@@

\caption{\label{tab:demographic-table}Summary of demographic information}
\centering
\begin{tabular}[t]{llllr}
\toprule
Preferred pronoun & Education & Age group & Previous experience & Subject\\
\midrule
She & Diploma and Bachelor Degree & 18-24 & No & 17\\
He & Diploma and Bachelor Degree & 18-24 & Yes & 14\\
She & Diploma and Bachelor Degree & 18-24 & Yes & 12\\
He & Diploma and Bachelor Degree & 18-24 & No & 11\\
He & Diploma and Bachelor Degree & 25-39 & No & 10\\
She & Diploma and Bachelor Degree & 25-39 & No & 10\\
She & Diploma and Bachelor Degree & 25-39 & Yes & 9\\
He & High School or below & 18-24 & No & 8\\
She & High School or below & 18-24 & No & 7\\
He & High School or below & 18-24 & Yes & 6\\
He & Diploma and Bachelor Degree & 25-39 & Yes & 5\\
He & Masters Degree & 25-39 & No & 5\\
He & Masters Degree & 25-39 & Yes & 5\\
He & High School or below & 25-39 & No & 4\\
She & High School or below & 25-39 & No & 4\\
She & Masters Degree & 25-39 & Yes & 4\\
He & High School or below & 25-39 & Yes & 3\\
She & High School or below & 18-24 & Yes & 3\\
She & Masters Degree & 25-39 & No & 3\\
He & Honours Degree & 25-39 & No & 2\\
She & Diploma and Bachelor Degree & 40-54 & No & 2\\
She & Honours Degree & 18-24 & No & 2\\
He & High School or below & 40-54 & Yes & 1\\
He & High School or below & 55-64 & No & 1\\
He & Masters Degree & 40-54 & Yes & 1\\
He & Masters Degree & 55-64 & No & 1\\
Other & High School or below & 25-39 & No & 1\\
She & High School or below & 40-54 & No & 1\\
She & Honours Degree & 25-39 & No & 1\\
She & Honours Degree & 25-39 & Yes & 1\\
She & Masters Degree & 18-24 & No & 1\\
She & Masters Degree & 40-54 & No & 1\\
They & Diploma and Bachelor Degree & 18-24 & No & 1\\
They & Diploma and Bachelor Degree & 25-39 & No & 1\\
They & High School or below & 18-24 & No & 1\\
They & High School or below & 25-39 & No & 1\\
\bottomrule
\end{tabular}
\end{table}

@@

@@

\hypertarget{shpaes-of-non-linearity}{%
\subsubsection{Shpaes of non-linearity}\label{shpaes-of-non-linearity}}

@@

\hypertarget{reasons-for-making-selections}{%
\subsubsection{Reasons for making
selections}\label{reasons-for-making-selections}}

@@

>>>>

% see https://stackoverflow.com/a/47122900

% Pandoc citation processing

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\begin{quote}
\emph{``Since all models are wrong the scientist must be alert to what
is importantly wrong \citep{box1976science}.''}
\end{quote}

Diagnosing a model is the key to determining whether there is anything
importantly wrong. For linear regression analysis, it is typical to
interrogate the residuals. Residuals summarise what is not captured by
the model, and thus provide the capacity to identify what might be
wrong. There are many ways that residuals could be assessed.

Residuals might be plotted, as a histogram or quantile-quantile plot to
examine the distribution. Using classical linear regression model as an
example, if the distribution is symmetric and unimodal, then it is
well-behaved. But if the distribution is skewed, bimodal, multimodal, or
even contains outliers, there is cause for concern. It is also possible
to conduct a goodness of fit test to examine the distribution of
residuals, such as Shapiro-Wilk Normality test
\citep{shapiro1965analysis}.

To examine the relationship between residuals, predicted values and each
of the explanatory variables, it is recommend to plot them on a scatter
plot with residuals on the y-axis. This type of plot is usually called
residual plot. In general, one will check whether there is any
noticeable pattern to determine if the model is well-behaved. However,
it is a very difficult task for a human judge, though to make a decision
that there's nothing there. It is especially common, particularly among
new data analysts to report patterns when an experienced data analyst
might quickly conclude that there are none. Generally, one looks for
departures from nothingness like nonlinear dependency or
heteroskedasticity. It is also possible to conduct a hypothesis test for
non-linear dependence \citep{ramsey_tests_1969}, and also use a
Breusch-Pagan test \citep{breusch_simple_1979} for heteroskedasticity.

There is an abundance of literature describing appropriate diagnostic
methods for linear regression: \citet{draper1998applied},
\citet{montgomery1982introduction}, \citet{belsley_regression_1980},
\citet{cook_applied_1999} and \citet{cook1982residuals}. The diligent
reader of these sage writings will also notice sentences that express
sentiments like \emph{based on their experience, statistical tests are
not widely used in regression diagnostics. The same or even larger
amount of information can be provided by diagnostic plots than the
corresponding tests in most empirical studies.} There is a common
guidance by experts that plots are the best for diagnosing model fits.

This is curious, and investigating why this might be common advice is
the subject of this paper. The paper is structured as follows. The next
background section describes the commonly used diagnostic plots, the
types of departures that one expects to detect, and describes a formal
process for reading residual plots, called visual inference, that can
avoid the concerns about subjectiveness of human readers. Section
\ref{experimental-design} describes the experimental setup to enable a
comparison between decision made by formal hypothesis testing, and how
humans would read diagnostic plots. The results are reported in Section
\ref{results}. We finish with a discussion on future work, in particular
how the responsibility for residual plot reading might be passed on to
computer vision.

@@

@@

In overall, there were a total of 3200 lineup evaluations made by 160
subjects in both experiment I and experiment II respectively, where 320
lineup evaluations were attention checks and were not used in the
following analysis.

The collated dataset is provided in \texttt{polynomials} and
\texttt{heter} of the \texttt{visage} \texttt{R} package.

@@

\caption{\label{tab:demographic-table}Summary of demographic information}
\centering
\begin{tabular}[t]{llllr}
\toprule
Preferred pronoun & Education & Age group & Previous experience & Subject\\
\midrule
She & Diploma and Bachelor Degree & 18-24 & Yes & 26\\
She & Diploma and Bachelor Degree & 18-24 & No & 24\\
He & Diploma and Bachelor Degree & 18-24 & No & 22\\
He & High School or below & 18-24 & No & 22\\
She & Diploma and Bachelor Degree & 25-39 & No & 21\\
He & Diploma and Bachelor Degree & 18-24 & Yes & 19\\
He & Diploma and Bachelor Degree & 25-39 & No & 19\\
She & High School or below & 18-24 & No & 19\\
He & High School or below & 18-24 & Yes & 15\\
She & Diploma and Bachelor Degree & 25-39 & Yes & 15\\
He & Diploma and Bachelor Degree & 25-39 & Yes & 12\\
She & High School or below & 18-24 & Yes & 10\\
He & Masters Degree & 25-39 & No & 8\\
He & Masters Degree & 25-39 & Yes & 8\\
He & High School or below & 25-39 & No & 7\\
She & Masters Degree & 25-39 & Yes & 7\\
She & High School or below & 25-39 & No & 6\\
He & High School or below & 25-39 & Yes & 5\\
She & Honours Degree & 25-39 & No & 5\\
She & Honours Degree & 25-39 & Yes & 4\\
She & Masters Degree & 25-39 & No & 4\\
He & Honours Degree & 25-39 & Yes & 3\\
She & Honours Degree & 18-24 & No & 3\\
He & Diploma and Bachelor Degree & 40-54 & Yes & 2\\
He & Diploma and Bachelor Degree & 55-64 & No & 2\\
He & Honours Degree & 25-39 & No & 2\\
He & Masters Degree & 40-54 & No & 2\\
She & Diploma and Bachelor Degree & 40-54 & No & 2\\
She & High School or below & 25-39 & Yes & 2\\
She & Masters Degree & 18-24 & No & 2\\
They & Diploma and Bachelor Degree & 18-24 & No & 2\\
They & Diploma and Bachelor Degree & 25-39 & No & 2\\
He & Diploma and Bachelor Degree & 40-54 & No & 1\\
He & High School or below & 40-54 & Yes & 1\\
He & High School or below & 55-64 & No & 1\\
He & Honours Degree & 18-24 & No & 1\\
He & Honours Degree & 18-24 & Yes & 1\\
He & Honours Degree & 40-54 & Yes & 1\\
He & Masters Degree & 40-54 & Yes & 1\\
He & Masters Degree & 55-64 & No & 1\\
Other & Diploma and Bachelor Degree & 18-24 & No & 1\\
Other & High School or below & 25-39 & No & 1\\
She & Diploma and Bachelor Degree & 40-54 & Yes & 1\\
She & High School or below & 40-54 & No & 1\\
She & High School or below & 55-64 & Yes & 1\\
She & Honours Degree & 40-54 & Yes & 1\\
She & Masters Degree & 40-54 & No & 1\\
They & High School or below & 18-24 & No & 1\\
They & High School or below & 18-24 & Yes & 1\\
They & High School or below & 25-39 & No & 1\\
\bottomrule
\end{tabular}
\end{table}

@@

\hypertarget{experiment-i}{%
\subsubsection{Experiment I}\label{experiment-i}}

@@

\hypertarget{shpaes-of-non-linearity}{%
\paragraph{Shpaes of non-linearity}\label{shpaes-of-non-linearity}}

@@

\hypertarget{reasons-for-making-selections}{%
\paragraph{Reasons for making
selections}\label{reasons-for-making-selections}}

@@

\hypertarget{experiment-ii}{%
\subsubsection{Experiment II}\label{experiment-ii}}
